Estimating duration in the presence of misclassification

### 2024-02-05 cis-imperfect-testing.tex
edits/comments from Overleaf

To assess the sensitivity of the procedure to a time-varying test sensitivity, it will be useful to simulate from a model with a varying test sensitivity as a function of days since the start of the infection episode.
Through consideration of \cref{imperf-test:fig:bounding-cis-sensitivity}, and the interaction of false negatives due to viral load in \cref{E-ATACCC}'s results, I propose the following model.
\begin{equation}
  p_\text{sens}(t) = \begin{cases}
    0.9 - \frac{0.9-0.5}{50}t &t \leq 50 \\
    0.5 &t > 50
  \end{cases}
  \label{imperf-test:eq:variable-test-sensitivity}
\end{equation}
where $t$ is the number of days since infection.
Therefore, this model consists of a linear decline in test sensitivity, before a constant minimum point.
Adapting the simulation study to incorporate this test sensitivity is much more similar to the data (see purple histogram in \cref{imperf-test:fig:sim-single-pos}).
This model does not require low test sensitivities early in the infection, maintaining compatibility with the results in \cref{E-ATACCC}.

@@

First, if $e_i \leq r_i^{(e)}$, meaning that the test at $r_i^{(e)}+1$ is a true negative and the end of the episode is interval-censored as in the previous chapter.
In this case, the test at $r_i^{(e)}$ is a true negative, as are all other tests not in $\sched'_i$.
These occur with probability 1, by the assumption of no false positives.
The tests in $\sched_i'$ are either true positives or false negatives.
\begin{align}
&p(\vec{y_i'}, e_i \leq r_i^{(e)} | b_i, p_\text{sens}, \theta) \\
&= p(\vec{y_i'}, l_i^{(e)} \leq e_i \leq r_i^{(e)} | t_i, b_i, p_\text{sens}, \theta) \\ % &\text{as no false positives}
&= p(\vec{y_i'} \mid l_i^{(e)} \leq e_i \leq r_i^{(e)}, t_i, b_i, p_\text{sens}, \theta) p(l_i^{(e)} \leq e_i \leq r_i^{(e)} | t_i, b_i, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_i(t)} (1 - p_\text{sens})^{(1 - y_i(t))} \right) \left( S_\theta(l_i^{(e)} - b_i - 1) - S_\theta(r_i^{(e)} - b_i - 1) \right)
\label{imperf-test:eq:ll-ei-lt-ri}
\end{align}

@@

@@

For each simulated dataset, I estimate the survival function using the model proposed in \cref{imperf-test:sec:modelling} and assuming $\psens$ equals 0.6, 0.8, or 1.0.
If the dataset was simulated using a constant $\psens$, and that value of $\psens$ matches the value assumed during inference, I refer to $\psens$ as being correctly specified; otherwise I refer to it as misspecified.
However, there is still some misspecification of the model to false negatives owing to the simplifying assumptions made.
The amount that the simplifying assumptions are violated increases as $\psens$ decreases.

@@

The results when a varying test sensitivity (\cref{imperf-test:eq:variable-test-sensitivity}) is used for the simulation are similar to a constant 0.8 test sensitivity (see \cref{imperf-test:fig:variable-test-sensitivity}).
This suggests that the simplified model, with constant test sensitivity, is sufficient for recovering the true survival time.
Therefore, I will apply this model to the real CIS data in the next section.
\begin{figure}
  % \makebox[\textwidth][c]{
    \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-variable-sensitivity}
  \caption[Simulation study results with varying test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a variable test sensitivity.
    Each panel shows the results of performing inference with a different assumed value for the test sensitivity.
    The posterior estimate using a model with a constant test sensitivity of 0.8 (middle panel) is similar to the true value (black line).
  }
  \label{imperf-test:fig:variable-test-sensitivity}
\end{figure}

>>>>

To assess the sensitivity of the procedure to a time-varying test sensitivity, it will be useful to simulate from a model with a varying test sensitivity as a function of days since the start of the infection episode.
Through consideration of \cref{imperf-test:fig:bounding-cis-sensitivity}, and the interaction of false negatives due to viral load in \cref{E-ATACCC}'s results, I propose the following model.
\begin{equation}
  p_\text{sens}(t) = \begin{cases}
    0.9 - \frac{0.9-0.5}{50}t &t \leq 50 \\
    0.5 &t > 50
  \end{cases}
  \label{imperf-test:eq:variable-test-sensitivity}
\end{equation}
where $t$ is the number of days since infection.
Therefore, this model consists of a linear decline in test sensitivity, before a constant minimum point.

This could be incorporated into the simulation study be generating test results at ...
Adapting the simulation study to incorporate this test sensitivity is much more similar to the data (see purple histogram in \cref{imperf-test:fig:sim-single-pos}).
This model does not require low test sensitivities early in the infection, maintaining compatibility with the results in \cref{E-ATACCC}.

@@

First, if $e_i \leq r_i^{(e)}$, meaning that the test at $r_i^{(e)}+1$ is a true negative and the end of the episode is interval-censored as in the previous chapter.
In this case, the test at $r_i^{(e)} + 1$ is a true negative, as are all other tests not in $\sched'_i$.
These occur with probability 1, by the assumption of no false positives.
The tests in $\sched_i'$ are either true positives or false negatives.
\begin{align}
&p(\vec{y_i'}, e_i \leq r_i^{(e)} | b_i, p_\text{sens}, \theta) \\
&= p(\vec{y_i'}, l_i^{(e)} \leq e_i \leq r_i^{(e)} | t_i, b_i, p_\text{sens}, \theta) \\ % &\text{as no false positives}
&= p(\vec{y_i'} \mid l_i^{(e)} \leq e_i \leq r_i^{(e)}, t_i, b_i, p_\text{sens}, \theta) p(l_i^{(e)} \leq e_i \leq r_i^{(e)} | t_i, b_i, p_\text{sens}, \theta) \\
&= \left( \prod_{t \in \sched'_i} p_\text{sens}^{y_i(t)} (1 - p_\text{sens})^{(1 - y_i(t))} \right) \left( S_\theta(l_i^{(e)} - b_i - 1) - S_\theta(r_i^{(e)} - b_i - 1) \right)
\label{imperf-test:eq:ll-ei-lt-ri}
\end{align}

@@

Constant terms => proprtional to simpler thing

@@

For each simulated dataset, I estimate the survival function using the likelihood proposed in \cref{imperf-test:sec:modelling} and assuming $\psens$ equals 0.6, 0.8, or 1.0.
If the dataset was simulated using a constant $\psens$, and that value of $\psens$ matches the value assumed during inference, I refer to $\psens$ as being correctly specified; otherwise I refer to it as misspecified.
However, there is still some misspecification of the model to false negatives owing to the simplifying assumptions made.
The amount that the simplifying assumptions are violated increases as $\psens$ decreases.

@@

The results when a varying test sensitivity (\cref{imperf-test:eq:variable-test-sensitivity}) is used for the simulation are similar to a constant 0.8 test sensitivity (see \cref{imperf-test:fig:variable-test-sensitivity}).
This suggests that the simplified model, with constant test sensitivity, is sufficient for recovering the true survival time.
Therefore, I will apply this model to the real CIS data in the next section.
\begin{figure}
  % \makebox[\textwidth][c]{
    \includegraphics[width=\textwidth]{cis-imperfect-testing/sim-variable-sensitivity}
  \caption[Simulation study results with varying test sensitivity]{%
    Posterior (median and 95\% credible interval) of the survival time for the simulation study with a variable test sensitivity.
    Each panel shows the results of performing inference with a different assumed value for the test sensitivity.
    The posterior estimate using a model with a constant test sensitivity of 0.8 (B) is similar to the true value (black line).
  }
  \label{imperf-test:fig:variable-test-sensitivity}
\end{figure}

### 2024-02-05 transmission.tex
edits/comments from Overleaf

\todo[inline]{Rewrite the intro: eg see Dani's Overleaf comments}

In this chapter, I use the preferred estimate of the duration of RT-PCR positivity (see \cref{E-imperf-test:sec:conclusion}) to estimate the transmission of SARS-CoV-2 in England.
This estimation is based only on the CIS data (see \cref{transmission:sec:data} for further details).
I will show that this data is sufficient to estimate the parameters of the transmission process, suggesting surveys of this type are sufficient for surveillance of an epidemic.

@@

\todo[inline]{Consider moving all of this section into chapter 2, where CIS is introduced, and spell it out a bit more in that context}

@@

In MRP, a Bayesian multilevel logistic regression model is first fit to predict RT-PCR positivity from demographic covariates.
Then, the posterior distribution is used to predict the prevalence in each stratum.
Finally, the prevalence is poststratified (as explained later) to the CIS target population and subpopulations of interest.
I modify this process by transforming the posterior distribution of prevalence in each stratum to a posterior distribution of incidence before poststratifying.
% I fit a modification model to estimate prevalence, then transform the posterior distribution of prevalence in each stratum to a posterior distribution of incidence.
% Finally, I poststratify the posterior distribution of incidence to the CIS target population and subpopulations of interest.

@@

The model was fit separately and independently for each region, implicitly adding all interactions between region and the other variables.
This step assumes that the estimates for each region can be considered independent, which is reasonable because the regions are large with many tests conducted within each.

@@

There are two places where this misspecification could occur.
The first is the survival matrix, $\matr{S}$.
For example, \cref{E-imperf-test} showed that it is sensitive to assumptions made.
The second is the specification of the statistical model used to estimate prevalence, for example the form of the smoothing used.

@@

The incidence proportion was highest among the 11--15 and 16--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 16--25 age group, all age groups show similar trends (\ie parallel lines on a log scale).
This is the pattern predicted by an age-stratified, mechanistic transmission model (such as the one introduced in \cref{E-SEIR:sec:structured-populations}), if behavioural changes are independent of age group, and there is negligible impact of immunity.
The incidence proportion in the 16--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts})
Many other explanations are possible, and further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using backcalculation by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

@@

The transmission part of the model follows \textcite{birrellRealtime}.
A novel approach (\cref{SEIR:sec:observation}) links the transmission model to the data.
In the mechanistic modelling literature, this part of the model is referred to as the observation model; however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.

@@

% Here, I extend the system of ODEs to include compartments representing RT-PCR positive individuals.
% The proportion of RT-PCR positive individuals in each strata is then linked to the CIS using a Beta-Binomial likelihood, with the modelled proportion that are RT-PCR positive as the mean of the Beta distribution.
One approach would be to convolve the incidence based on the transmission model with the RT-PCR duration distribution.
Here, the daily incidence proportion for each stratum is by solving the system of ODEs governing the transmission model, and then prevalence calculated as using \cref{E-inc-prev:eq:EPt}.
However, this approach is computationally expensive, likely prohibitively so if performed within a MCMC loop.

@@

I link the modelled proportion of individuals that are RT-PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{transmission:sec:data}).
If on each day $t$, each strata $i$ is observed to have $y_{it}$ positives out of $n_{it}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \sum_{k=1}^l \bar{p}_{kat}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where $\bar{p}_{kat} = \int_{t}^{t+1} p_{ka}(s) ds$ is the mean occupancy in compartment $p_k$ and strata $a$ over day $t$ and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.

@@

    \node[draw, thick, inner sep=0.3cm, fit=(P1) (P3), fill=blue!10,opacity=0.2] {};
\end{tikzpicture}
}
  \caption[SEIR model]{SEIR model used for the analysis in this section. The shaded box represents RT-PCR-positive individuals. Shown for a single stratum.}
  \label{SEIR:fig:full-model}
\end{figure}

@@

During such an exponential growth phase, all components of the $\vec{e}$, $\vec{i}$ and $\vec{p}$ compartments grow at the same growth rate, $\psi$.
This follows directly from the ODEs if $\vec{s}$ is constant. 
The growth rate $\psi$ is easily identifiable from the data, and therefore a parameterisation in terms of $\phi$ should be more computationally efficient.
If this is the initial exponential phase, when $\vec{s} \approx 1$, then the basic reproduction number for a SEIR model is linked to this growth rate by~\autocites{birrellBayesian}{wearingAppropriate}:
\begin{align}
    \R = \frac{\psi \left( \frac{\psi}{2\sigma} + 1 \right)^2}{\gamma \left(1 - \frac{1}{\left(\frac{\psi}{2 \gamma} + 1 \right)^2} \right)} \label{SEIR:eq:rtoR}.
\end{align}

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow.
This is a common issue with random walk MCMC (see \cref{E-inc-prev:sec:MCMC}).
This is similar when considering coverage of the derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \includegraphics{SEIR/sim/coverage}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=0.9\paperwidth]{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for incidence (top) and prevalence (bottom); see text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

@@

Overall the model fits the data well (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
There are a few features of the data which are not captured by the model and the posterior estimate appears to be a compromise with the prevalence too low in one age group and too high in another.
For example, in the North East region, the 11--16 age group has its prevalence overestimated in October but the 16--25 age group has its prevalence underestimated.
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_young}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (younger ages)]{%
        Goodness-of-fit to prevalence data by region and age for the younger three age groups (see \cref{SEIR:fig:prev-old} for the others).
        The central ribbon and line shows the posterior median and 95\% CrI estimate for the proportion of the relevant strata RT-PCR positive on each day.
        The lighter outer ribbon is the 95\% posterior predictive interval for the proportion of test results that are positive each week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev-young}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_old}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (older ages)]{%
        Goodness-of-fit prevalence for the older three age groups (see \cref{SEIR:fig:prev-young} for the other age groups and details).
    }
    \label{SEIR:fig:prev-old}
\end{figure}

@@

These hypotheses for the mechanism causing changes in $\beta_w$ are correlated across the country.
This can explain that similar patterns emerge for $\beta_w$ in different regions (see \cref{SEIR:fig:beta-walk}), despite these being estimated independently (\ie there is no model structure which favours correlated values).
These correlations could be exploited to improve the precision of estimates with a more complex model which borrows information across regions.

@@

\Cref{transmission:fig:compare-regions} compares the regional incidence estimates between the two models in this chapter.
Generally the estimates agree, with the SEIR model having much narrower credible intervals.
The largest discrepancy is that the backcalculation model estimates a second peak in mid January in several regions (most clearly in London, North West, and the East of England).
However, except in London, the credible intervals still overlap.
This is a specific case of a larger pattern: the SEIR model produces smoother estimates of incidence than the backcalculation model.;
\begin{figure}
    \centering \includegraphics{transmission/compare-regions}
    \caption[Comparing each models estimate of regional incidence.]{%
        Incidence proportion in England and each region estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the backcalculation results.
    }
    \label{transmission:fig:compare-regions}
\end{figure}

@@

Many of these issues are clearly demonstrated by the estimates for the North East (see \cref{transmission:fig:compare-NE}).
I previously observed the quality of the SEIR model's fit is variable between different age groups (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
This is because the model is not flexible enough to capture the true age distribution of infections.
The overestimation of prevalence in the 11--16 age group and underestimation in the 16--25 age group is reflected when comparing incidence estimates to those from the backcalculation model, which have reduced bias.
\begin{figure}
    \centering \includegraphics{transmission/compare-NE}
    \caption[Comparing each models estimate of North East incidence by age.]{%
        Incidence proportion in each age group in the North East estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly RT-PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the backcalculation results.
    }
    \label{transmission:fig:compare-NE}
\end{figure}

>>>>

\todo[inline]{Intro probably needs rewriting, with novelty claims and a bit more context. Revisit once intro chapters written.}

In this chapter, I use the preferred estimate of the duration of PCR positivity (see \cref{E-imperf-test:sec:conclusion}) to estimate the transmission of SARS-CoV-2 in England.
This estimation is based only on the CIS data (see \cref{transmission:sec:data} for further details).
I will show that this data is sufficient to estimate the parameters of the transmission process, suggesting surveys of this type are sufficient for surveillance of an epidemic.

@@

\todo[inline]{This is now mostly in chapter 2, double-check there's nothing missed

@@

In MRP, a Bayesian multilevel logistic regression model is first fit to predict PCR positivity from demographic covariates.
Then, the posterior distribution is used to predict the prevalence in each stratum.
Finally, the prevalence is poststratified (as explained later) to the CIS target population and subpopulations of interest.
I modify this process by transforming the posterior distribution of prevalence in each stratum to a posterior distribution of incidence before poststratifying.
% I fit a modification model to estimate prevalence, then transform the posterior distribution of prevalence in each stratum to a posterior distribution of incidence.
% Finally, I poststratify the posterior distribution of incidence to the CIS target population and subpopulations of interest.

The deterministic relationship between incidence proportion and prevalence means that either can be inferred, with the other then calculated.
Denote the quantity to be inferred at time $t$ is strata $i$ as $\pi_{it}$.
Note that for any $i$ and $t$, the number of positives will be small, often fewer than 5; therefore some form of borrowing between time and strata is required.
The statistical model used for inference, including the borrowing of information, should have the following properties.
\begin{itemize}
    \item The smoothing model should penalise towards exponential growth.
        Penalising towards exponential growth is a natural formulation because mechanistic models predict exponential growth over short periods of time.
        Furthermore, such penalisation has previously been shown to improve estimates of prevalence by balancing over and under-fitting~\autocite{ealesAppropriately}.
\end{itemize}

@@

The model was fitted separately and independently for each region, implicitly adding all interactions between region and the other variables.
This step assumes that the estimates for each region can be considered independent, which is reasonable because the regions are large with many tests conducted within each.

@@

There are two places where this misspecification could occur.
The first is the survival matrix, $\matr{S}$.
For example, \cref{E-imperf-test} showed that it is sensitive to assumptions made, such as the assumed $\psens$ and ....
The second is the specification of the statistical model used to estimate prevalence, for example the form of the smoothing used.

@@

The incidence proportion was highest among the 11--15 and 16--25 age groups throughout the period (see \cref{transmission:fig:backcalc-ages}).
Except the 16--25 age group, all age groups show similar trends (\ie parallel lines on a log scale).
The distribution of infections among age groups remaining constant over short periods of time is the pattern predicted by standard age-stratified, mechanistic transmission model (such as the one introduced in \cref{E-SEIR:sec:structured-populations}) in equilibrium, if behavioural changes are independent of age group, and there is negligible impact of immunity.
The incidence proportion in the 16--25 age group decreases faster than other age groups during November.
There are many plausible explanations for this effect.
The higher incidence in this age group prior to November means it could be an immunity effect.
Alternatively, it could be a consequence of schools remaining open in the November lockdown; 20--24 year olds rarely interact with school-aged children (see \cref{E-SEIR:fig:age-contacts})
Many other explanations are possible, and further work is required to determine causality.
\begin{figure}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{transmission/backcalc-ages}}
    \caption[Incidence estimated using backcalculation by age group]{%
        Incidence proportion for each age group, on both natural and log scales.
    }
    \label{transmission:fig:backcalc-ages}
\end{figure}

@@

The transmission part of the model follows \textcite{birrellRealtime}.
A novel approach (\cref{SEIR:sec:observation}) links the transmission model to prevalence data (here from the CIS, but the approach is generalisable).
In the mechanistic modelling literature, this part of the model is referred to as the observation model; however, in the terminology of \cref{E-inc-prev}, it includes both the prevalence process and the observation process.

@@

% Here, I extend the system of ODEs to include compartments representing PCR positive individuals.
% The proportion of PCR positive individuals in each strata is then linked to the CIS using a Beta-Binomial likelihood, with the modelled proportion that are PCR positive as the mean of the Beta distribution.
One approach would be to convolve the incidence based on the transmission model with the PCR duration distribution.
Here, the daily incidence proportion for each stratum is by solving the system of ODEs governing the transmission model, and then prevalence calculated as using \cref{E-inc-prev:eq:EPt}.
However, this approach is computationally expensive, likely prohibitively so if performed within a MCMC loop.

@@

I link the modelled proportion of individuals that are PCR positive to the data using a beta-binomial likelihood (as discussed in \cref{transmission:sec:data}).
If on each day $t$, each strata $i$ is observed to have $y_{it}$ positives out of $n_{it}$ tests, then the likelihood is:
\begin{align}
    \prod_{a,t} \BB (y_{at} \mid n_{at}, \sum_{k=1}^l \bar{p}_{kat}, \rho)
    \label{SEIR:eq:likelihood}
\end{align}
where $\bar{p}_{kat} = \int_{t}^{t+1} p_{ka}(s) ds$ is the mean occupancy in compartment $p_k$ and strata $a$ over day $t$ and $\rho$ is a nuisance parameter controlling the overdispersion of the beta-binomial distribution.

@@

    \node[draw, thick, inner sep=0.3cm, fit=(P1) (P3), fill=blue!10,opacity=0.2] {};
\end{tikzpicture}
}
  \caption[SEIR model]{SEIR model used for the analysis in this section. The shaded box represents PCR-positive individuals. Shown for a single stratum.}
  \label{SEIR:fig:full-model}
\end{figure}

@@

During such an exponential growth phase, all components of the $\vec{e}$, $\vec{i}$ and $\vec{p}$ compartments grow at the same growth rate, $\psi$.
This follows directly from the ODEs if $\vec{s}$ is constant. 
The growth rate $\psi$ is easily identifiable from the data, and therefore a parameterisation in terms of $\psi$ should be more computationally efficient.
If this is the initial exponential phase, when $\vec{s} \approx 1$, then the basic reproduction number for a SEIR model is linked to this growth rate by~\autocites{birrellBayesian}{wearingAppropriate}:
\begin{align}
    \R = \frac{\psi \left( \frac{\psi}{2\sigma} + 1 \right)^2}{\gamma \left(1 - \frac{1}{\left(\frac{\psi}{2 \gamma} + 1 \right)^2} \right)} \label{SEIR:eq:rtoR}.
\end{align}

@@

The majority of the confidence intervals contain the nominal probability (see \cref{SEIR:fig:sim-coverage}).
There is a tendency for the coverage to be slightly lower than the nominal probability, suggesting that the credible intervals are slightly too narrow.
This is a common issue with random walk MCMC (see \cref{E-inc-prev:sec:MCMC}).
This is similar when considering coverage of the derived quantities incidence and prevalence (see \cref{SEIR:fig:sim-inc-prev}).
\begin{figure}
    \includegraphics{SEIR/sim/coverage}
    \caption[Coverage of simulation study]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for each parameter; see text for full definition.
        Vertical lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-coverage}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \todo[inline]{Add (A) incidence to the top panel and (B) prevalence to the bottom panel}
    \makebox[\textwidth][c]{\includegraphics[width=0.9\paperwidth]{SEIR/sim/predictive_coverage}}
    \caption[Coverage of simulation study (derived quantities)]{%
        Point estimate and 95\% confidence interval for coverage of the 50\%, 75\%, and 95\% central credible intervals for incidence (top) and prevalence (bottom); see text for full definition.
        Horizontal lines show the nominal coverage.
    }
    \label{SEIR:fig:sim-inc-prev}
\end{figure}

@@

Overall the model fits the data well (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
There are a few features of the data which are not captured by the model and the posterior estimate appears to be a compromise with the prevalence too low in one age group and too high in another.
For example, in the North East region, the 11--16 age group has its prevalence overestimated in October but the 16--25 age group has its prevalence underestimated.
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_young}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (younger ages)]{%
        Goodness-of-fit to prevalence data by region and age for the younger three age groups (see \cref{SEIR:fig:prev-old} for the others).
        The central ribbon and line shows the posterior median and 95\% CrI estimate for the proportion of the relevant strata PCR positive on each day.
        The lighter outer ribbon is the 95\% posterior predictive interval for the proportion of test results that are positive each week.
        The dots show the observed prevalence in CIS, aggregated by week.
        For a well-calibrated model, 95\% of the dots would be within the outer ribbon.
    }
    \label{SEIR:fig:prev-young}
\end{figure}
\begin{figure}
    \thisfloatpagestyle{empty}
    \vspace{-3cm}
    \makebox[\textwidth][c]{\includegraphics[width=.9\paperwidth]{SEIR/CIS/prev_old}}
    \captionsetup{width=0.8\paperwidth}
    \captionof{figure}[SEIR prevalence goodness-of-fit (older ages)]{%
        Goodness-of-fit prevalence for the older three age groups (see \cref{SEIR:fig:prev-young} for the other age groups and details).
    }
    \label{SEIR:fig:prev-old}
\end{figure}

@@

These hypotheses for the mechanism causing changes in $\beta_w$ are correlated across the country.
This can explain that similar patterns emerge for $\beta_w$ in different regions (see \cref{SEIR:fig:beta-walk}), despite these being estimated independently (\ie there is no model structure which favours correlated values).
These correlations could be exploited to improve the precision of estimates with a more complex model which borrows information across regions, \eg a hierarchical model.

@@

\Cref{transmission:fig:compare-regions} compares the regional incidence estimates between the two models in this chapter.
Generally the estimates agree, with the SEIR model having much narrower credible intervals.
The largest discrepancy is that the backcalculation model estimates a second peak in mid January in several regions (most clearly in London, North West, and the East of England).
However, except in London, the credible intervals still overlap.
This is a specific case of a larger pattern: the SEIR model produces smoother estimates of incidence than the backcalculation model.;
\begin{figure}
    \centering \includegraphics{transmission/compare-regions}
    \caption[Comparing each models estimate of regional incidence.]{%
        Incidence proportion in England and each region estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the backcalculation results.
    }
    \label{transmission:fig:compare-regions}
\end{figure}

@@

Many of these issues are clearly demonstrated by the estimates for the North East (see \cref{transmission:fig:compare-NE}).
I previously observed the quality of the SEIR model's fit is variable between different age groups (see \cref{SEIR:fig:prev-young,SEIR:fig:prev-old}).
This is because the model is not flexible enough to capture the true age distribution of infections.
The overestimation of prevalence in the 11--16 age group and underestimation in the 16--25 age group is reflected when comparing incidence estimates to those from the backcalculation model, which have reduced bias.
\begin{figure}
    \centering \includegraphics{transmission/compare-NE}
    \caption[Comparing each models estimate of North East incidence by age.]{%
        Incidence proportion in each age group in the North East estimated by the two models discussed in this chapter.
        Posterior median and 95\% CrI shown.
        For the SEIR model, newly PCR positive (entrances into the P1 compartment), rather than incidence (entrances into the E compartment) are shown for comparability with the backcalculation results.
    }
    \label{transmission:fig:compare-NE}
\end{figure}
